Image projections along an axis

### 2006-02-12 article/Article.tex
add sigma projection
darcs-hash:20060211120717-2fc9d-d551ef7f2866c3802acb6209b295cb2667e56cc4.gz
enhance the article
darcs-hash:20060212220651-2fc9d-eb0a1bf23226000511562e56af95bd9d85b27bf6.gz

% The title should be descriptive enough for people to be able to find
% the relevant document. 
\title{Image projections along an axe}

@@

\section{Introduction}
ITK already provide a class to project an image along an axe:
\doxygen{AccumulateImageFilter}. However, this class is limited to sum
projection and mean projection.

@@

The base class, ProjectionImageFilter, is mainly a copy of
AccumulateImageFilter, with some enhancement:
\begin{itemize}
  \item projection is delegated to a functor
  \item report its progress
\end{itemize}
Different projection types are provided, with specialized classes:
\begin{itemize}
  \item maximum projection (MaximumProjectionImageFilter)
  \item minimum projection (MinimumProjectionImageFilter)
  \item mean projection (MeanProjectionImageFilter)
  \item median projection (MedianProjectionImageFilter)
  \item sum projection (SumProjectionImageFilter)
\end{itemize}
Other projections types can easily be created just by providing a new functor.

@@

\section{Implemeting a new projection method}
As an example, here is a filter which implement a square sum projection along a
axe. This example may be useful in case of signed input type.

@@

namespace Function {
\end{verbatim} \normalsize
Now, create the class which do the real work: receive all the pixel values and
give the result at the end. The class is created in the itk::Function names pace
to avoid putting to much things in itk name space, and is templated over the
type of the input and the output pixel.
On object of this class is created for each pixel in the output image.
\small \begin{verbatim}
template <class TInputPixel, class TOuputPixel>
class SquareSumAccumulator
{
public:
\end{verbatim} \normalsize
The constructor must take one parameter: the number of pixels to accumulate
during the projection. In that case, this parameter is not useful, so it is not
saved in an attribute of the class. The sum is initialized to zero.

The destructor is empty - there is nothing to do.
\small \begin{verbatim}
  SquareSumAccumulator( unsigned long nb )
    {
    m_SquareSum = NumericTraits< TOuputPixel >::Zero;
    }
  ~SquareSumAccumulator(){}
\end{verbatim} \normalsize
The operator() is called for each pixel. In this method, the square sum is
contructed, pixel after pixel. The method is declared inline for best
performance.
\small \begin{verbatim}
  inline TInputPixel operator()( const TInputPixel &input )
    {
    m_SquareSum = m_SquareSum + ( input * input );
    }
\end{verbatim} \normalsize
GetValue() is called to set the value of the result image. In that example, the
sum is just return, but more complex computation can be done here (see the
MedianProjectionImageFilter for an example). ProjectionImageFilter cast the
output of this method to the output image's pixel type.
\small \begin{verbatim}
  inline TOuputPixel GetValue()
    {
    return m_SquareSum;
    }
\end{verbatim} \normalsize
The internal attribute which accumulate the square sum.
\small \begin{verbatim}
  TOuputPixel m_SquareSum;
};
} // end namespace Function
\end{verbatim} \normalsize
The class is declared. It inherits from ProjectionImageFilter, and declares the
usual types and methods.
\small \begin{verbatim}
template <class TInputImage, class TOutputImage>
class ITK_EXPORT SquareSumProjectionImageFilter :
    public
    ProjectionImageFilter<TInputImage, TOutputImage,
      Function::SquareSumAccumulator< typename TInputImage::PixelType,
        typename TOutputImage::PixelType > >
{
public:
  typedef SquareSumProjectionImageFilter Self;
  typedef ProjectionImageFilter<TInputImage, TOutputImage, 
    Function::SquareSumAccumulator< typename TInputImage::PixelType,
      typename TOutputImage::PixelType > > Superclass;

@@

\section{Examples}
Figure~\ref{maximum0}, Figure~\ref{maximum1}, Figure~\ref{maximum},
Figure~\ref{minimum}, Figure~\ref{mean} and Figure~\ref{median} show the result
of different projection on different axes. The result of
SumProjectionImageFilter in similar to the result of MeanProjectionImageFilter,
but with a higher range of values, and so is not shown here.

\begin{figure}[htbp]
\centering
\includegraphics[scale=0.5]{maximum0}
\caption{MaximumProjectionImageFilter, projection along axe x.\label{maximum0}}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics[scale=0.5]{maximum1}
\caption{MaximumProjectionImageFilter, projection along axe y.\label{maximum1}}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics[scale=0.75]{maximum}
\caption{MaximumProjectionImageFilter, projection along axe z.\label{maximum}}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics[scale=0.75]{minimum}
\caption{MinimumProjectionImageFilter, projection along axe z.\label{minimum}}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics[scale=0.75]{mean}
\caption{MeanProjectionImageFilter, projection along axe z.\label{mean}}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics[scale=0.75]{median}
\caption{MedianProjectionImageFilter, projection along axe z\label{median}}
\end{figure}

@@

\small \begin{verbatim}
#include "itkImageFileReader.h"
#include "itkImageFileWriter.h"
#include "itkCommand.h"
#include "itkSimpleFilterWatcher.h"

@@

int main(int, char * argv[])
{
  int axe = atoi(argv[1]);

@@

  typedef itk::MaximumProjectionImageFilter< IType, IType > FilterType;
  FilterType::Pointer filter = FilterType::New();
  filter->SetInput( reader->GetOutput() );
  filter->SetAxe( axe );

@@

  IType::SizeType size;
  for(int i=0; i<=3; i++) {
    if(i == axe) {
      size[i] = 0;
    } else {
      size[i] = inputSize[i];
     }
  }
  IType::IndexType idx;
  idx.Fill(0);

@@

>>>>

% The title should be descriptive enough for people to be able to find
% the relevant document. 
\title{Image projections along an axis}

@@

\section{Introduction}
ITK already provide a class to project an image along an axe:
\doxygen{AccumulateImageFilter}. However, this class is limited to sum
and mean projections.

@@

The base class, \verb$ProjectionImageFilter$, is mainly a copy of
\verb$AccumulateImageFilter$, with some enhancement:
\begin{itemize}
  \item projection is delegated to a functor
  \item report its progress
  \item thread support
\end{itemize}
Different projection types are provided, with specialized classes:
\begin{itemize}
  \item maximum projection (\verb$MaximumProjectionImageFilter$)
  \item minimum projection (\verb$MinimumProjectionImageFilter$)
  \item mean projection (\verb$MeanProjectionImageFilter$)
  \item median projection (\verb$MedianProjectionImageFilter$)
  \item sum projection (\verb$SumProjectionImageFilter$)
  \item sigma projection (\verb$SigmaProjectionImageFilter$)
  \item binary projection (\verb$BinaryProjectionImageFilter$)
\end{itemize}
Other projections types can easily be created just by providing a new functor.

The projection can be easily threaded by splitting the required region. The
Only important thing is to take care to not split the region on the projection
dimension. 
The default split done by \verb$ImageSource$ if fully usable without any changes
because it is done on the output region, and that region have a size of 1 on
the projection dimension, so the split will be done on anothe dimesion.

@@

\section{Implemeting a new projection method}
As an example, here is a filter which implement a square sum projection along a
axis. This example may be useful in case of signed input type.

@@

namespace Function {
\end{verbatim} \normalsize
Now, create the class which do the real work: receive all the pixel values and
give the result at the end. The class is created in the \verb$itk::Function$ names pace
to avoid putting too much things in itk name space, and is templated over the
type of the input and the output pixel.
One object of this class is created for each thread.
\small \begin{verbatim}
template <class TInputPixel, class TOuputPixel>
class SquareSumAccumulator
{
public:
\end{verbatim} \normalsize
The constructor must take one parameter: the number of pixels to accumulate
during the projection. In that case, this parameter is not useful, so it is not
saved in an attribute of the class.

If some custom parameters need to be passed to the accumulator, the \verb$NewAccumulator()$
method can be overloaded. See \verb$BinaryProjectionImageFilter$'s code for an
example.

The destructor is empty - there is nothing to do.
\small \begin{verbatim}
  SquareSumAccumulator( unsigned long size ) {}
  ~SquareSumAccumulator(){}
\end{verbatim} \normalsize
\verb$Init()$ is called before accumulating a new set of pixel
to initialize the accumulator. Here, the sum is initialized to zero.
\small \begin{verbatim}
  void Init()
    {
    m_SquareSum = NumericTraits< TOuputPixel >::Zero;
    }
\end{verbatim} \normalsize
The \verb$operator()$ is called for all pixels. In this method, the square sum is
contructed, pixel after pixel. The method is declared inline for best
performance.
\small \begin{verbatim}
  inline TOuputPixel operator()( const TInputPixel &input )
    {
    m_SquareSum = m_SquareSum + vnl_math_sqr( (TOuputPixel) input );
    }
\end{verbatim} \normalsize
\verb$GetValue()$ is called to set the value of the result image. In that example, the
sum is just returned, but more complex computations can be done here (see the
\verb$MedianProjectionImageFilter$'s code for an example). \verb$ProjectionImageFilter$ cast the
output of this method to the output image's pixel type.
\small \begin{verbatim}
  inline TOuputPixel GetValue()
    {
    return m_SquareSum;
    }
\end{verbatim} \normalsize
The internal attribute which accumulate the square sum.
\small \begin{verbatim}
  TOuputPixel m_SquareSum;
};
} // end namespace Function
\end{verbatim} \normalsize
The class is declared. It inherits from ProjectionImageFilter, and declares the
usual types and methods.
\small \begin{verbatim}
template <class TInputImage, class TOutputImage>
class ITK_EXPORT SquareSumProjectionImageFilter :
    public
    ProjectionImageFilter<TInputImage, TOutputImage,
      Function::SquareSumAccumulator< typename TInputImage::PixelType,
        typename TOutputImage::PixelType > >
{
public:
  typedef SquareSumProjectionImageFilter Self;
  typedef ProjectionImageFilter<TInputImage, TOutputImage, 
    Function::SquareSumAccumulator< typename TInputImage::PixelType,
      typename TOutputImage::PixelType > > Superclass;

@@

\section{Examples}
Figure~\ref{maximum0}, Figure~\ref{maximum1}, Figure~\ref{maximum},
Figure~\ref{minimum}, Figure~\ref{mean}, Figure~\ref{median},
Figure~\ref{sigma} and Figure~\ref{binary100} show the result
of different projection on different axes. The result of
SumProjectionImageFilter in similar to the result of MeanProjectionImageFilter,
but with a higher range of values, and so is not shown here.

\begin{figure}[htbp]
\centering
\includegraphics[scale=0.5]{maximum0}
\caption{MaximumProjectionImageFilter, projection along axis x.\label{maximum0}}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics[scale=0.5]{maximum1}
\caption{MaximumProjectionImageFilter, projection along axis y.\label{maximum1}}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics[scale=0.75]{maximum}
\caption{MaximumProjectionImageFilter, projection along axis z.\label{maximum}}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics[scale=0.75]{minimum}
\caption{MinimumProjectionImageFilter, projection along axis z.\label{minimum}}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics[scale=0.75]{mean}
\caption{MeanProjectionImageFilter, projection along axis z.\label{mean}}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics[scale=0.75]{median}
\caption{MedianProjectionImageFilter, projection along axis z\label{median}}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics[scale=0.75]{sigma}
\caption{SigmaProjectionImageFilter, projection along axis z\label{sigma}}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics[scale=0.75]{binary100}
\caption{BinaryProjectionImageFilter, projection along axis z, with foreground value = 100, and background value = 255\label{binary100}}
\end{figure}

@@

\small \begin{verbatim}
#include "itkSimpleFilterWatcher.h"

@@

int main(int, char * argv[])
{
  int dim = atoi(argv[1]);

@@

  typedef itk::MaximumProjectionImageFilter< IType, IType > FilterType;
  FilterType::Pointer filter = FilterType::New();
  filter->SetInput( reader->GetOutput() );
  filter->SetProjectionDimension( dim );
  // to be sure that the result is ok with several threads, even on a single
  // proc computer
  filter->SetNumberOfThreads( 2 );

@@

  IType::SizeType size;
  for(int i=0; i<=3; i++) {
    if(i == dim) {
      size[i] = 0;
    } else {
      size[i] = inputSize[i];
     }
  }
  IType::IndexType idx;
  idx.Fill(0);

@@

  std::cout << "size: " << size << std::endl;
  std::cout << "index: " << idx << std::endl;

  extract->Update();
  std::cout << "output size: " << extract->GetOutput()->GetLargestPossibleRegion().GetSize() << std::endl;
